GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer

### 2020-04-14 paper/paper_main.tex
some editing throughout the paper draft

% GravExplain
\documentclass[prb,preprint]{revtex4-1} 
% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}

% Note that AJP uses the same style as Phys. Rev. B (prb).
% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
% We look forward to receiving your submission to AJP.
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{float}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{hyperref}
% \usepackage{siunitx}
\usepackage{textcomp}

@@

@@

\author{James W. Gardner}
\email{u6069809@anu.edu.au}
\affiliation{College of Science, Australian National University, Acton, ACT, 2601, Australia}

\author{Hannah Middleton$^{(i)}$}
\email{hannah.middleton@unimelb.edu.au}
\author{Andrew Melatos$^{(i)}$}
\email{amelatos@unimelb.edu.au}
\author{Robin Evans$^{(ii)}$}
\email{robinje@unimelb.edu.au}
\author{William Moran$^{(ii)}$}
\email{wmoran@unimelb.edu.au}
\affiliation{School of Physics$^{(i)}$ and Department of Electrical and Electronic Engineering$^{(ii)}$, University of Melbourne, Parkville, VIC, 3010, Australia}
\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}

@@

\section{Observing a single tone}
\label{sec:single_tone}
% the webcam method
% how to analyse video, python OpenCV2

As mentioned in Section~\ref{sec:introduction}, continuous gravitational wave searches look for near monochromatic signals. The audio analogue of this is a simple sinusoidal tone; a constant note.

We measured the intensity at a point on the screen with a commercial USB webcam placed as to view the screen as in Figure~\ref{fig:ifo_schematic_webcam}. This webcam samples at 30Hz, which limits the spectral content of signals we can observe to below 15Hz, the Nyquist frequency.
% See later Section~\ref{sec:photodiode} where we raise this limit into audible frequencies (20 Hz to 20kHz) with a photodiode.
While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam. Then, at an off-centre pixel (shown as a red dot in Figure~\ref{fig:interference_pattern}), we took the green-channel of the video as approximate to the total intensity, as the laser is monochromatic green light (at 532nm).

We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. The two tones played together sound similar and don’t produce any noticeable beating.
To quantify the precision, we recorded a signal from an injected tone of 2.09Hz. From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\mathrm{Hz}$, a result within two standard deviations. This showed that the experiment observes a single tone confidently.

\begin{figure}[h]
	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the amplitude 17.5 peak was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at 0Hz, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
	\label{fig:webcam_spectrum}
\end{figure}

>>>>

% GravExplain
\documentclass[prb,reprint]{revtex4-1} 
% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}

% Note that AJP uses the same style as Phys. Rev. B (prb).
% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
% We look forward to receiving your submission to AJP.
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{float}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{hyperref}
% \usepackage{siunitx}
\usepackage{textcomp}
\usepackage{subfiles}

@@

\author{James W. Gardner}
\email{u6069809@anu.edu.au}
\affiliation{%
College of Science, Australian National University, Acton, ACT, 2601, Australia
}%

\author{Changrong Liu}
\email{}
\affiliation{%
Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
}

\author{Hannah Middleton}
\email{hannah.middleton@unimelb.edu.au}
\affiliation{%
 School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
}
\affiliation{%
OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
}

\author{Andrew Melatos}
\email{amelatos@unimelb.edu.au}
\affiliation{%
 School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
}
\affiliation{%
OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
}

\author{Robin Evans}
\email{robinje@unimelb.edu.au}
\affiliation{%
Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
}
\affiliation{%
OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
}

\author{William Moran}
\email{wmoran@unimelb.edu.au}
\affiliation{%
Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
}
\affiliation{
}

\author{\han{also invite the people who helped us with interferometer design to be authors??}}
\email{wmoran@unimelb.edu.au}
\affiliation{%
Adelaide, Adelaide-OzGrav
}

\begin{abstract}
Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy data. 
We demonstrate these techniques by observing sound with a table-top Michelson interferometer. 
In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings. 
However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
%Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data. 
%We demonstrate these techniques by observing sound with a table-top Michelson interferometer. 
%In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
%We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings. 
%However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
\subfile{ifo-introduction.tex}

\section{Table-top gravitational wave science}
\label{sec:ifo}
\subfile{ifo-tableTopGWs.tex}

\section{Observing a single tone}
\label{sec:single_tone}
\subfile{ifo-singleTone.tex}

\section{Observing a wandering tone}
\label{sec:viterbi_wandering}
\subfile{ifo-wanderingTone.tex}

\section{Playing back complex audio}
\label{sec:optical_microphone}
\subfile{ifo-complexAudio.tex}

\section{Future work}
\label{sec:future_work}
\subfile{ifo-futureWork.tex}

\section{Conclusions}
\label{sec:conclusions}
\subfile{ifo-conclusions.tex}

\begin{acknowledgments}
The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer. 
This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
Travel during the project was supported by the Australian National University PhB Science program.

\newpage
\appendix
\subfile{ifo-appendix.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\author{James W. Gardner}
\email{u6069809@anu.edu.au}
\affiliation{%
College of Science, Australian National University, Acton, ACT, 2601, Australia
}%

\author{Hannah Middleton}
\email{hannah.middleton@unimelb.edu.au}
\affiliation{%
 School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
}
\affiliation{%
OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
}

\author{Andrew Melatos}
\email{amelatos@unimelb.edu.au}
\affiliation{%
 School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
}
\affiliation{%
OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
}

\author{Robin Evans}
\email{robinje@unimelb.edu.au}
\affiliation{%
Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
}
\affiliation{%
OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
}

\author{William Moran}
\email{wmoran@unimelb.edu.au}
\affiliation{%
Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
}

@@
